Positional surprisal (1-3grams) in CHILDES corpora

### 2019-05-13 surprisal/bnc&switchboard.Rmd
set up wiki_relative_surprisals

swc <- sw %>% filter(length %in% c(10, 20, 30))
switch_surprisals <- map_dfr(1:nrow(swc), 
        ~bind_cols(select(slice(swc, .x), length, word_order), 
                               tidyboot_mean(data = unnest(slice(swc, .x)), 
                                             column = s, nboot = 100)))

>>>>

swc <- sw %>% filter(length %in% c(10, 20, 30))
switch_surprisals <- map_dfr(1:nrow(swc), 
        ~bind_cols(select(slice(swc, .x), length, word_order), 
                               tidyboot_mean(data = unnest(slice(swc, .x)), 
                                             column = s, nboot = 1000)))

### 2019-05-13 surprisal/childes_surprisal.Rmd
set up wiki_relative_surprisals

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(gtools)
library(here)

@@

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
zhou_utterances <- get_utterances(corpus = "ZhouDinner")
```

```{r}
adult_utterances <- zhou_utterances %>%
  filter(speaker_role == "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, stem) %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())
```

```{r unigrams}
adult_unigrams <- adult_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

@@

adult_surprisals %>%
  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

>>>>

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(gtools)
library(here)
library(ggthemes)

@@

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
shiro_utterances <- get_utterances(corpus = "Shiro")
```

```{r old unigrams}
adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

adult_unigrams <- adult_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

@@

adult_surprisals %>%
  filter(length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

```{r unigrams}
get_unigrams <- function(utterances, speaker_role, stem_or_gloss) {

  if (tolower(speaker_role) == "child") {
    utts <- utterances %>%
      filter(speaker_role == "Target_Child") 
  } else {
    utts <- utterances %>%
      filter(speaker_role != "Target_Child")   
  }

  if (tolower(stem_or_gloss) == "stem") {
    utts <- utts %>%
      mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
      mutate(utterance_id = 1:n()) %>%
      unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
      group_by(utterance_id) %>%
      mutate(word_order = 1:n())

  } else {
    utts <- utts %>%
      mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
      mutate(utterance_id = 1:n()) %>%
      unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
      group_by(utterance_id) %>%
      mutate(word_order = 1:n())
  }

  unigrams <- utts %>%
    group_by(word) %>%
    count() %>%
    ungroup() %>%
    mutate(p = n / sum(n)) %>%
    mutate(s = -log(p)) %>%
    select(word, s)

  surprisals <- utts %>%
    left_join(unigrams) %>%
    group_by(length, word_order) %>%
    summarise(s = mean(s)) %>%
    ungroup()
}

prov_sups <- get_unigrams(prov_utterances, "not child", "stem") %>% 
  mutate(speaker_role = "adult") %>%
  bind_rows(mutate(get_unigrams(prov_utterances, "child", "stem"), speaker_role = "child")) %>%
  mutate(language = "English")

shiro_sups <- get_unigrams(shiro_utterances, "not child", "gloss") %>% 
  mutate(speaker_role = "adult") %>%
  bind_rows(mutate(get_unigrams(shiro_utterances, "child", "gloss"), speaker_role = "child")) %>%
  mutate(language = "Spanish")

zhou_sups <- get_unigrams(zhou_utterances, "not child", "stem") %>% 
  mutate(speaker_role = "adult") %>%
  bind_rows(mutate(get_unigrams(zhou_utterances, "child", "stem"), speaker_role = "child")) %>%
  mutate(language = "Mandarin")

all_sups <- prov_sups %>%
  bind_rows(shiro_sups, zhou_sups)

label_data <- data_frame(length = 9, language = "Mandarin", word_order = 6,
                         speaker_role = c("adult", "child"), s = c(7,9.5))

all_sups %>% 
  filter(length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s,
                        color = speaker_role,
                        label = speaker_role)) +
    facet_grid(language ~ length, scales = "free") + 
    xlab("Word position") +
    ylab("Mean surprisal") +
    geom_line(size = .5) +
    theme_classic(base_size = 10) + 
    scale_color_ptol() +
    theme(legend.position = "nones") + 
    geom_text(data = label_data)
